---
title: "MSDS-6372-Project2"
author: "Holmes, Karki, Webb"
date: '11/12/2022'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Set-up
__Libraries__
```{r load-packages, include=FALSE}
## Source - https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/
# Package names
packages <- c("aplore3", "tidyverse")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())

if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

```{r echo=FALSE}
packages
```
__Data Import__
```{r}
df <- glow_bonemed
```

__EDA__
```{r}
```


### Objective 1
Display the ability to perform EDA and build a logistic regression model for interpretation purposes.
__Simple Logistic Regression__
```{r}

```

### Objective 2
With a simple logistic regression model as a baseline, perform additional competing models to improve on prediction performance metrics.
__Complex Logistic Regression__
```{r}
```

__LDA/QDA__
```{r}
```

__KNN__
```{r}
```







